---
title: "Movie Syntax"
author: "Thomas Sostarics"
date: "8/14/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Setup

```{r setup, include=FALSE}
library(reticulate)
library(tidyverse)
library(RSQLite)
library(sqldf)
library(dbplyr)
library(hrbrthemes)
library(palettetown)
knitr::opts_chunk$set(echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)
```


# Create database

Set up database to query later. For some reason I ***cannot*** get `data.tsv`
to load with `dbWriteTable()` when I give it the local filename. For some
reason it only works when I read it into memory first, which is **definitely**
not ideal. I think it has something to do with the single quote characters in
some of the movie titles. `akas.tsv` doesn't have the same issue though.

```{r create db}
con <- dbConnect(SQLite(), dbname = "movies_db")

large_data <- read_tsv('data.tsv')
dbWriteTable(con, name="movies", value=large_data, overwrite=T,
             row.names=FALSE, header=TRUE, sep = "\t")
rm(large_data) # Free up a chunk of memory

dbWriteTable(con, name="akas", value="akas.tsv", overwrite=T,
             row.names=FALSE, header=TRUE, sep = "\t")
```

# SQL Query

Run our query, essentially we want:
  - Only movies
  - Nothing with missing genres or release dates
  - No News, Documentary, Biography, or History movies
  - Preferential ranking of Horror > Adventure > Comedy > Action > Drama >>> Other
  when a movie has more than one genre.
  - Only movies with English titles
  
The cases short circuit, so Horror will take precedence over Adventure and so on.

```{sql, connection=con, output.var="queried"}
SELECT DISTINCT
  d.tconst,
  d.genres,
  d.primaryTitle AS title,
  d.startYear AS year,
  CASE 
    WHEN genres LIKE '%Horror%' THEN  'Horror'
    WHEN genres LIKE '%Adventure%' THEN 'Adventure'
    WHEN genres LIKE '%Action%' THEN 'Action'
    WHEN genres = 'Comedy' THEN 'Comedy'
    WHEN genres = 'Drama' THEN 'Drama'
    ELSE 'Other'
  END AS 'dominant_genre'
FROM movies AS d
LEFT JOIN akas AS a
  ON d.tconst = a.titleID
WHERE 
  a.region        = 'US'
  AND d.genres   <> '\\N'
  AND d.titleType = 'movie'
  AND startYear  <> '\\N'
  AND d.primaryTitle = d.originalTitle
  AND genres NOT LIKE '%News%'
  AND genres NOT LIKE '%Documentary%'
  AND genres NOT LIKE '%Biography%'
  AND genres NOT LIKE '%History%'
  AND dominant_genre <> 'Other';
```

# Run POS tagging in Python

spaCy builds a syntactic tree, so I add a carrier sentence to prevent
things like 'Red Planet' being incorrectly labeled as 'PROPN PROPN'
instead of 'ADJ NOUN'.

```{python}
import pandas
import numpy
import nltk
import spacy
import pickle
nlp = spacy.load('en')
titles = [nlp('this movie is titled ' + title.lower()) for title in r.queried.title]

print(len(titles))

titles_tags = []

for title in titles:
  title_pos = ' '.join([w.pos_ for w in title[4:]]) #[4:] removes the carrier
  titles_tags.append(title_pos)
```

# Post processing in R

Here we add in the POS tags we got from python, then group and plot the data.

```{r, warning=F}
# Combine our POS tagging with our original movie data
queried$tag <-  py$titles_tags
queried$decade <- queried$year %/% 10 * 10

# Get the counts of each tag form for every genre
form_counts <- 
  queried %>% 
  filter(dominant_genre!='Other', year>=1980, year<2020) %>% 
  group_by(dominant_genre, tag, decade) %>% 
  summarize(count = n()) 

# Get the total number that each tag form appears
tag_totals <- 
  form_counts %>% 
  group_by(tag,decade) %>% 
  summarize(decade_total = sum(count)) %>% 
  group_by(tag) %>% 
  mutate(tag_total = sum(decade_total))

# Add in the tag totals and genre counts 
form_counts <-
  form_counts %>% 
  left_join(tag_totals, by=c('tag', 'decade')) %>% 
  filter(!is.na(tag_total),
         tag_total>700) %>% 
  group_by(dominant_genre, decade) %>% 
  mutate(genre_total = sum(count)) %>% 
  ungroup()

# Calculate proportions and opacity
form_probabilities <- 
  form_counts %>% 
  mutate(proportion = count/genre_total) %>%
  group_by(tag, decade) %>% 
  mutate(is_top = ifelse(proportion==max(proportion),1,.35), #1=opaque
         decade = paste0(as.character(decade),"s")) 

# Plot data
form_probabilities %>% 
  ggplot(aes(x=dominant_genre, y=proportion, fill=tag)) +
  geom_bar(stat='identity', color='grey16', aes(alpha=I(is_top))) +
  hrbrthemes::theme_ipsum() +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_poke(pokemon=232,spread=6) +
  ylab('Proportion of Movies using a Syntactic Form') +
  xlab('Genre') +
  ggtitle("Movie Title Syntactic Form Across Genres") +
  labs(fill = 'Syntax') +
  facet_wrap(~decade)+
  scale_y_continuous(labels = scales::percent)+
  labs(caption = 
         'While overall proportions of movie title syntax have not varied much across genres over the past 40 years, and movies named "Noun"
          (Silence; Eternity; Dolls) consistently account for 25-35% of movies, Horror movies have consistently had a slightly higher
         proportion of movies named "The Noun" (The Shining; The Conjuring; The Thing) and Action movies have grown to 
        have the highest proportion of movies with titles that are just "Proper Noun" (Steampunk; Icarus; Jacksonville).
       (Opaque bars highlight which genre has the highest proportion for that syntactic form)
       Data from IMDB
       @tsostaricsling')
```

